Longitudinal Data Cleaning - A Case of the NLSY79 Data

### 2021-02-22 report/report.tex
Add part of IDA.

\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## tidy the gender and race variables}
\NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
\NormalTok{         SAMPLE_RACE_78SCRN,}
\NormalTok{         SAMPLE_SEX_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{gender =}\NormalTok{ SAMPLE_SEX_}\DecValTok{1979}\NormalTok{,}
         \DataTypeTok{race =}\NormalTok{ SAMPLE_RACE_78SCRN)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# tidy the grade }
\NormalTok{demog_education <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{as_tibble}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# in 2018, the variable's name is Q3-4_2018, instead of HGC_2018}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{HGC_2018 =} \StringTok{`}\DataTypeTok{Q3-4_2018}\StringTok{`}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
         \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HGCREV"}\NormalTok{),}
         \StringTok{"HGC_2012"}\NormalTok{,}
         \StringTok{"HGC_2014"}\NormalTok{,}
         \StringTok{"HGC_2016"}\NormalTok{,}
         \StringTok{"HGC_2018"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
               \DataTypeTok{names_to =} \StringTok{"var"}\NormalTok{,}
               \DataTypeTok{values_to =} \StringTok{"grade"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{separate}\NormalTok{(}\StringTok{"var"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"var"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(grade)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{var)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## getting the highest year of completed education ever}
\NormalTok{highest_year <-}\StringTok{ }\NormalTok{demog_education }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{hgc_i =} \KeywordTok{max}\NormalTok{(grade)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\NormalTok{grade) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(year }\OperatorTok{==}\StringTok{ }\KeywordTok{first}\NormalTok{(year)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{yr_hgc =}\NormalTok{ year) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{, yr_hgc, hgc_i) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{'hgc'}\NormalTok{ =}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\StringTok{"NONE"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{, }\StringTok{"1ST GRADE"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{2}\NormalTok{, }\StringTok{"2ND GRADE"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{, }\StringTok{"3RD GRADE"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{>=}\StringTok{ }\DecValTok{4} \OperatorTok{&}\StringTok{ }\NormalTok{hgc_i }\OperatorTok{<=}\StringTok{ }\DecValTok{12}\NormalTok{, }\KeywordTok{paste0}\NormalTok{(hgc_i,}\StringTok{"TH GRADE"}\NormalTok{),}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{13}\NormalTok{, }\StringTok{"1ST YEAR COL"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{14}\NormalTok{, }\StringTok{"2ND YEAR COL"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{15}\NormalTok{, }\StringTok{"3RD YEAR COL"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{95}\NormalTok{, }\StringTok{"UNGRADED"}\NormalTok{, }
                        \KeywordTok{paste0}\NormalTok{((hgc_i }\OperatorTok{-}\StringTok{ }\DecValTok{12}\NormalTok{), }\StringTok{"TH YEAR COL"}\NormalTok{)))))))))))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# make a list for years where we used the "QES-52A"}
\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
\CommentTok{#function to get the hour of work}
\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
\NormalTok{   temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
  \ControlFlowTok{else}\NormalTok{\{}
\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
\NormalTok{    temp }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
\NormalTok{\}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
\NormalTok{  new_data_qnames }\OperatorTok{%>%}
\StringTok{     }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
\NormalTok{\}}
\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{   rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
\NormalTok{\}}
\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
\CommentTok{# join hours and rates variable}
\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all, }
\NormalTok{                         hours_all, }
                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
\NormalTok{                                rate_per_hour))}
\KeywordTok{head}\NormalTok{(hours_wages)}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{mean\_hourly\_wage} and \texttt{full\_demographic} data are
then joined. We also filtered the data to only have the cohort who
completed the education up to 12th grade and participated at least five
rounds in the survey.

@@

\hypertarget{robust-linear-model-for-influential-values-treatment}{%
\subsection{Robust Linear Model for Influential Values
Treatment}\label{robust-linear-model-for-influential-values-treatment}}

\hypertarget{conclusion}{%
\section*{Conclusion}\label{conclusion}}
\addcontentsline{toc}{section}{Conclusion}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-stringr}{}%
Wickham, Hadley. 2019. \emph{Stringr: Simple, Consistent Wrappers for
Common String Operations}.
\url{https://CRAN.R-project.org/package=stringr}.

>>>>

\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## tidy the gender and race variables}
\NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
\NormalTok{         SAMPLE_RACE_78SCRN,}
\NormalTok{         SAMPLE_SEX_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{gender =}\NormalTok{ SAMPLE_SEX_}\DecValTok{1979}\NormalTok{,}
         \DataTypeTok{race =}\NormalTok{ SAMPLE_RACE_78SCRN)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# tidy the grade }
\NormalTok{demog_education <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{as_tibble}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# in 2018, the variable's name is Q3-4_2018, instead of HGC_2018}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{HGC_2018 =} \StringTok{`}\DataTypeTok{Q3-4_2018}\StringTok{`}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
         \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HGCREV"}\NormalTok{),}
         \StringTok{"HGC_2012"}\NormalTok{,}
         \StringTok{"HGC_2014"}\NormalTok{,}
         \StringTok{"HGC_2016"}\NormalTok{,}
         \StringTok{"HGC_2018"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
               \DataTypeTok{names_to =} \StringTok{"var"}\NormalTok{,}
               \DataTypeTok{values_to =} \StringTok{"grade"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{separate}\NormalTok{(}\StringTok{"var"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"var"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(grade)) }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{var)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## getting the highest year of completed education ever}
\NormalTok{highest_year <-}\StringTok{ }\NormalTok{demog_education }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{hgc_i =} \KeywordTok{max}\NormalTok{(grade)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\NormalTok{grade) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(year }\OperatorTok{==}\StringTok{ }\KeywordTok{first}\NormalTok{(year)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{yr_hgc =}\NormalTok{ year) }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{, yr_hgc, hgc_i) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{'hgc'}\NormalTok{ =}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\StringTok{"NONE"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{, }\StringTok{"1ST GRADE"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{2}\NormalTok{, }\StringTok{"2ND GRADE"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{, }\StringTok{"3RD GRADE"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{>=}\StringTok{ }\DecValTok{4} \OperatorTok{&}\StringTok{ }\NormalTok{hgc_i }\OperatorTok{<=}\StringTok{ }\DecValTok{12}\NormalTok{, }\KeywordTok{paste0}\NormalTok{(hgc_i,}\StringTok{"TH GRADE"}\NormalTok{),}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{13}\NormalTok{, }\StringTok{"1ST YEAR COL"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{14}\NormalTok{, }\StringTok{"2ND YEAR COL"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{15}\NormalTok{, }\StringTok{"3RD YEAR COL"}\NormalTok{,}
                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{95}\NormalTok{, }\StringTok{"UNGRADED"}\NormalTok{, }
                        \KeywordTok{paste0}\NormalTok{((hgc_i }\OperatorTok{-}\StringTok{ }\DecValTok{12}\NormalTok{), }\StringTok{"TH YEAR COL"}\NormalTok{)))))))))))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# make a list for years where we used the "QES-52A"}
\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
\CommentTok{#function to get the hour of work}
\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
\NormalTok{   temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
  \ControlFlowTok{else}\NormalTok{\{}
\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
\NormalTok{    temp }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
\NormalTok{\}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
\NormalTok{  new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
\NormalTok{\}}
\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{   rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
\NormalTok{\}}
\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
\CommentTok{# join hours and rates variable}
\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all, }
\NormalTok{                         hours_all, }
                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
\NormalTok{                                rate_per_hour))}
\KeywordTok{head}\NormalTok{(hours_wages)}
\end{Highlighting}
\end{Shaded}

@@

The \texttt{mean\_hourly\_wage} and \texttt{full\_demographic} data are
then joined. We also filtered the data to only have the cohort who
completed the education up to 12th grade and participated at least five
rounds in the survey and save it to an object called
\texttt{wages\_demog\_hs}.

@@

According to Huebner, Vach, and Cessie (2016), Initial Data Analysis
(IDA) is the step of inspecting and screening the data after being
collected to ensure that the data is clean, valid, and ready to be
deployed in the later formal statistical analysis. Moreover, Chatfield
(1985) argued that the two main objectives of IDA is data description,
which is to assess the structure and the quality of the data; and model
formulation without any formal statistical inference.

In this paper, we conducted an IDA or a preliminary data analysis to
assess the consistency of the data with the cohort information that is
provided by the NLSY. In addition, we also aimed to find the anomaly in
the wages values using this approach. We mainly used graphical summary
to do the IDA using \texttt{ggplot2}(Wickham 2016) and \texttt{brolgar}
(Tierney, Cook, and Prvan 2020).

As stated previously, the respondents' ages ranged from 12 to 22 when
first interviewed in 1979. Hence, we would like to validate whether all
of the respondents were in this range. For the demographic data
validation, we used the \texttt{full\_demographic} i.e.~the data with
the survey years 1979 sample.

\begin{figure}
\centering
\includegraphics{report_files/figure-latex/valid-age-1.pdf}
\caption{Subjects' Age Distribution}
\end{figure}

\hypertarget{robust-linear-model-for-influential-values-treatment}{%
\subsubsection{Robust Linear Model for Influential Values
Treatment}\label{robust-linear-model-for-influential-values-treatment}}

\hypertarget{summary}{%
\section*{Summary}\label{summary}}
\addcontentsline{toc}{section}{Summary}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal
of the Royal Statistical Society. Series A. General} 148 (3): 214--53.

\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
Huebner, PhD, Marianne, Dr rer. nat Vach Werner, and PhD le Cessie
Saskia. 2016. ``A Systematic Approach to Initial Data Analysis Is Good
Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular
Surgery} 151 (1): 25--27.

\leavevmode\hypertarget{ref-brolgar}{}%
Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse
over Longitudinal Data Graphically and Analytically in R}.
\url{https://github.com/njtierney/brolgar}.

\leavevmode\hypertarget{ref-ggplot2}{}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\leavevmode\hypertarget{ref-stringr}{}%
---------. 2019. \emph{Stringr: Simple, Consistent Wrappers for Common
String Operations}. \url{https://CRAN.R-project.org/package=stringr}.
